---
title: "Calculate Heights from Las files"
author: "Michael Schmidt"
date: "1/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries
```{r}
library(tidyverse)
library(raster)
library(lidR)
library(sf)
```

# Make Heights from .las file

Method one.  This work pretty well, but takes a long time. 
```{r}
file_path <- "las_test/las_files/SM_BARKELEWDRAWNW_0301_DAT.177.dtm.las"

las<-readLAS(file_path)
```

## Clip so it runs a bit faster
```{r}
clipped_test<-st_read("las_test/clipped_tests/clip_shape.shp")

crop_las<-lasclip(las, clipped_test)
plot(crop_las)
str(crop_las)
```


## Trying another method for Calculating heights
```{r}
source("R/height_calc_funcs.R")
dz<-make_heights(crop_las)

hist(crop_las_ground_rig_1$height@data@values)

chm$dsm@data@values

plot(crop_las_ground_rig_2)
plot(crop_las_ground_rig_1)


```


Conbine functions together

```{r}
r_test<-add_heights("las_test/las_files/SM_BARKELEWDRAWNW_0301_DAT.177.dtm.las", "las_test/test_tiles/SM_BarkelewDrawNW_0301_180611.tif")


```

I tested this.  But at the moment it returns too many layers:
names      :      layer.1,      layer.2,      layer.3,      layer.4, SM_Barkel//1_180611.1, SM_Barkel//1_180611.2, SM_Barkel//1_180611.3, SM_Barkel//1_180611.4 
```{r}
start_time <- Sys.time()
combine_test<-add_heights("las_test/las_files/SM_BARKELEWDRAWNW_0301_DAT.177.dtm.las", "las_test/test_tiles/SM_BarkelewDrawNW_0301_180611.tif")
end_time <- Sys.time()
end_time-start_time

```


```{r}
test<-make_heights(file_path)

test2<-height_merge("las_test/test_tiles/SM_BarkelewDrawNW_0301_180611.tif", test)

tile<-brick("las_test/test_tiles/SM_BarkelewDrawNW_0301_180611.tif")

names(tile)<-c('band1', 'band2', 'band3', 'band4')
croped_tile<-test%>%
  crop(tile$band1)%>%
  mask(tile$band1)%>%
  stack(tile)
```

Testing add_heights function
```{r}
start_time <- Sys.time()
combine_test<-add_heights("las_test/las_files/SM_BASINSW_0103_DATE.118.dtm.las", "las_test/test_tiles/SM_BasinSW_0103_180611.tif")
end_time <- Sys.time()
end_time-start_time

start_time <- Sys.time()
combine_test<-add_heights("las_test/las_files/SM_BASINNE_0302_DATE.176.dtm.las", "las_test/test_tiles/SM_BasinNE_0302_180611.tif")
end_time <- Sys.time()
end_time-start_time

writeRaster(combine_test$height, "las_test/output/SM_BASINNE_0302_dz2.tif")
```

Testing csf() inputs for make_heights function.
```{r}
start_time <- Sys.time()
test_wt<-make_heights("las_test/las_files/SM_BASINSW_0103_DATE.118.dtm.las")
end_time <- Sys.time()
end_time-start_time

writeRaster(test_wt, "las_test/output/SM_BASINSW_0103_dz_sloop_smooth.tif")

start_time <- Sys.time()
raster_wt<-raster("las_test/test_tiles/SM_BasinSW_0103_180611.tif")
end_time <- Sys.time()
end_time-start_time

test_wt_rig_3<-make_heights("las_test/las_files/SM_BASINSW_0103_DATE.118.dtm.las")

writeRaster(test_wt, "las_test/output/SM_BASINSW_0103_dz_sloop_smooth_rig_3.tif")


```



