---
title: "Tile Checks"
author: "Michael Schmidt"
date: "3/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Check output compression
The Tiffs seem to already be as compressed as they can get. 
```{r}
library(raster)


tile_check<-raster("model_runs/Shadow_Class_MSAVI_No_Scale_V2/MODEL_shadow_model_and_MSAVI_V2_TILE_SM_BarkelewDrawSE_0401_180611.tif")

tile_check

writeRaster(tile_check, "output/compression_test/test_tile.tif")
writeRaster(tile_check, "output/compression_test/test_tile2.tif",options="COMPRESS=LZW")
writeRaster(tile_check, "output/compression_test/banana", datatype="INT2U" ,options="COMPRESS=LZW", overwrite=T)
```

## This is a test of the approach below
```{r}
file_paths<-list.files("model_runs/Shadow_Class_MSAVI_No_Scale_V2/", full.names=T, pattern=".tif$")

raster_list<-lapply(file_paths[1:3], raster)

raster_merge_list<-list()

raster_merge_list$filename<-raster_list

raster_list$overwrite<-FALSE

test_raster<-do.call(merge, raster_list)
plot(test_raster)
```
Sage = 5
```{r}
file_paths<-list.files("model_runs/Shadow_Class_MSAVI_No_Scale_V2/", full.names=T, pattern=".tif$")

time_start<-Sys.time()

raster_list<-lapply(file_paths, raster)

raster_list$overwrite<-FALSE

merged_raster<-do.call(merge, raster_list)

Sys.time()-time_start

writeRaster(merged_raster, "model_runs/Shadow_Class_MSAVI_No_Scale_V2/merged_MODEL_shadow_model_and_MSAVI_V2_non_compressed.tif")
```


## Working with moving window

### 1. 1 for sage 0 for everything else

I'm starting with a small tile as a test. 

```{r}
library(raster)
library(tidyverse)

rasterOptions(
  chunksize = 1e+09,
  maxmemory = 1e+10
)


tile_check<-raster("model_runs/Shadow_Class_MSAVI_No_Scale_V2/MODEL_shadow_model_and_MSAVI_V2_TILE_SM_BarkelewDrawSE_0401_180611.tif")

tile_check[tile_check %in% c(6, 4, 3, 2, 1)]<-0

tile_check[tile_check==5]<-1

sage_cells<-function(x){
  x[x %in% c(6, 4, 3, 2, 1)]<-0

  x[x==5]<-1
  
  return(x)
}

tile_check<-sage_cells(tile_check)

plot(tile_check)

ag_5m_sage<-aggregate(tile_check, fact = 50, fun=sum, na.rm=T)
ag_5m_sage%>%
  plot()

window_5m_3_by_3<-focal(ag_5m_sage, w=matrix(1/22500, nrow=3, ncol=3))
window_5m_5_by_5<-focal(ag_5m_sage, w=matrix(1/62500, nrow=5, ncol=5))

writeRaster(window_5m_3_by_3, "output/aggregate_test/window_5m_pixels_15m_window.tiff")
writeRaster(window_5m_5_by_5, "output/aggregate_test/window_5m_pixels_25m_window.tiff")

plot(window_5m_5_by_5)

```

## using reclassify so that we don't blow up the memory

```{r}
m<-c(1,0, 2,0, 3,0, 4,0, 5,1, 6,0)

rclmat<-matrix(m, ncol=2, byrow=T)

rc<-reclassify(tile_check, rclmat)

beginCluster(4, type='SOCK')
clusterR(tile_check, reclassify, args=list(rcl=rclmat))
endCluster()

plot(rc)

rc
```


```{r}
library(raster)
rasterOptions(tmpdir="output/raster_temp",
              chunksize = 1e+09,
              maxmemory = 1e+10)

dcb<-raster("model_runs/Shadow_Class_MSAVI_No_Scale_V2/merged_MODEL_shadow_model_and_MSAVI_V2_non_compressed.tif")

m<-c(1,0, 2,0, 3,0, 4,0, 5,1, 6,0)
rclmat<-matrix(m, ncol=2, byrow=T)

sage<-reclassify(dcb, rclmat)

writeRaster(sage, "model_runs/Shadow_Class_MSAVI_No_Scale_V2/products/sage_only.tif")

sage_only<-raster("model_runs/Shadow_Class_MSAVI_No_Scale_V2/products/sage_only.tif")

sage_only_5m<-aggregate(sage_only, fact = 50)


writeRaster(sage_only_5m, "model_runs/Shadow_Class_MSAVI_No_Scale_V2/products/sage_only_5m.tif")

plot(sage_only_5m)
```
## I failed to do everything above on the whole basin, so below we are going to do a tiled approach. 
```{r}

aggregate_single<-function(filepath,rclmat,fact, output_folder=FALSE){
  raster<-raster(filepath)
  reclassified_raster<-reclassify(raster, rclmat)
  
  if(output_folder==FALSE){
    aggregate_raster<-aggregate(reclassified_raster, fact=fact, fun=sum, na.rm=T)
    return(aggregate_raster)
  }else{
    aggregate(reclassified_raster, fact=fact, fun=sum, na.rm=T, filename=paste0(output_folder, "/5m_", basename(filepath)))
  }
}

fp<-"model_runs/Shadow_Class_MSAVI_No_Scale_V2/MODEL_shadow_model_and_MSAVI_V2_TILE_SM_BarkelewDrawSE_0401_180611.tif"

m<-c(1,0, 2,0, 3,0, 4,0, 5,1, 6,0)
matrix_stuff<-matrix(m, ncol=2, byrow=T)

test<-aggregate_single(fp, matrix_stuff, fact=50)

start_time<-Sys.time()
aggregate_single(fp, matrix_stuff, fact=50, output_folder="output/aggregate_test")
Sys.time()-start_time


plot(test)

```

## Make it Parallel
```{r}
file_paths<-list.files(
  "model_runs/Shadow_Class_MSAVI_No_Scale_V2", 
  full.names=TRUE, 
  pattern = ".tif$")

library(parallel)
library(doParallel)
n_cores<-detectCores(logical=TRUE)-2

cl<-makeCluster(n_cores)
registerDoParallel(cl)

clusterEvalQ(cl, library(raster))
clusterExport(cl, list('aggregate_single'))

parLapply(cl, 
          file_paths, 
          fun=aggregate_single, 
          rclmat=matrix_stuff, 
          fact=50,
          output_folder="model_runs/Shadow_Class_MSAVI_No_Scale_V2/products/sage_aggregate")

stopCluster(cl)
```

## merge th 5m rasters
```{r}
library(raster)
library(dplyr)
file_paths<-list.files("model_runs/Shadow_Class_MSAVI_No_Scale_V2/products/sage_aggregate", full.names=T, pattern=".tif$")

read_raster<-function(x){
  raster<-raster(x)
  names(raster)<-c("values")
  return(raster)
}

raster_list<-lapply(file_paths, read_raster)

new_raster_list<-raster_list[106:132]

new_raster_list$overlap<-FALSE
new_raster_list$tolerance<-2

start_time<-Sys.time()
merged_raster<-do.call(merge, new_raster_list)
Sys.time()-start_time

plot(merged_raster)

raster::merge(raster_list[[24]], raster_list[[25]], tolerance=0.5)%>%
  plot()

plot(raster_list[[25]])
plot(raster_list[[24]], add=T)
```
Everything above has failed.  We now try gdal utilities. 
```{r}
library(gdalUtils)

gdalbuildvrt(gdalfile="model_runs/Shadow_Class_MSAVI_No_Scale_V2/*.tif",
             output.vrt = "model_runs/Shadow_Class_MSAVI_No_Scale_V2/products/dev.vrt")


```



## merge and parallelizing a aggregate. 
```{r}
library(terra)
library(SpaDES)
tiles<-c("SM_GypsumGapNW_0102_180611.tif", "SM_GypsumGapNW_0103_180611.tif", "SM_GypsumGapNW_0104_180611.tif", "SM_GypsumGapNW_0202_180611.tif", "SM_GypsumGapNW_0203_180611.tif", "SM_GypsumGapNW_0204_180611.tif", "SM_NaturitaNWSW_0402_180611.tif", "SM_NaturitaNWSW_0403_180611.tif", "SM_NaturitaNWSW_0404_180611.tif")

read_raster<-function(x){
  r<-rast(paste0("model_runs/Shadow_Class_MSAVI_No_Scale_V2/MODEL_shadow_model_and_MSAVI_V2_TILE_", x))
  return(r)
}

raster_list<-lapply(tiles, read_raster)
raster_list[[9]]

raster_list$filename<-"model_runs/Shadow_Class_MSAVI_No_Scale_V2/products/merge_test.tif"

do.call(merge, raster_list)


test<-raster::raster("model_runs/Shadow_Class_MSAVI_No_Scale_V2/products/merge_test.tif")

raster::blockSize(test)

test_split<-splitRaster(test, nx=5, ny=5)

aggregate_single<-function(x,rclmat,fact, output_folder=FALSE){

  reclassified_raster<-reclassify(x, rclmat)
  
  if(output_folder==FALSE){
    aggregate_raster<-aggregate(reclassified_raster, fact=fact, fun=sum, na.rm=T)
    return(aggregate_raster)
  }else{
    aggregate(reclassified_raster, fact=fact, fun=sum, na.rm=T, filename=paste0(output_folder, "/5m_", basename(x@file@name)))
  }
}


m<-c(1,0, 2,0, 3,0, 4,0, 5,1, 6,0)
matrix_stuff<-matrix(m, ncol=2, byrow=T)

library(parallel)
library(doParallel)
n_cores<-detectCores(logical=TRUE)-2

cl<-makeCluster(n_cores)
registerDoParallel(cl)

clusterEvalQ(cl, library(raster))
clusterExport(cl, list('aggregate_single'))

parLapply(cl, 
          test_split, 
          fun=aggregate_single, 
          rclmat=matrix_stuff, 
          fact=50,
          output_folder="model_runs/Shadow_Class_MSAVI_No_Scale_V2/products/sage_aggregate2")

stopCluster(cl)

str(test_split[[1]])


basename(test_split[[1]]@file@name)

list_files_5m<-list.files("model_runs/Shadow_Class_MSAVI_No_Scale_V2/products/sage_aggregate2", full.names=T, pattern=".grd$")

files<-lapply(list_files_5m, raster::raster)

files$filename<-"model_runs/Shadow_Class_MSAVI_No_Scale_V2/products/merge_test4.tif"
files$tolerance<-1

do.call(raster::merge, files)


```



```{r}
library(raster)

tiles<-c("SM_GypsumGapNW_0102_180611.tif", "SM_GypsumGapNW_0103_180611.tif", "SM_GypsumGapNW_0104_180611.tif", "SM_GypsumGapNW_0202_180611.tif", "SM_GypsumGapNW_0203_180611.tif", "SM_GypsumGapNW_0204_180611.tif", "SM_NaturitaNWSW_0402_180611.tif", "SM_NaturitaNWSW_0403_180611.tif", "SM_NaturitaNWSW_0404_180611.tif")

read_raster<-function(x){
  r<-raster(paste0("model_runs/Shadow_Class_MSAVI_No_Scale_V2/products/sage_aggregate/5m_MODEL_shadow_model_and_MSAVI_V2_TILE_", x))
  return(r)
}

r<-lapply(tiles, read_raster)
r$tolerance<-0.5



t<-do.call(merge, r)

plot(t)
```

```{r}
library(raster)

files<-list.files("model_runs/Shadow_Class_MSAVI_No_Scale_V2/products/sage_aggregate/", full.names=T, pattern=".tif$")

files<-files[-177]
files<-files[-116]

r<-lapply(files, raster)

r$tolerance<-1



t<-do.call(merge, r)

plot(t)

writeRaster(t3, "model_runs/Shadow_Class_MSAVI_No_Scale_V2/products/sage_aggregate/merged/sage_merge_final.tif")

r2<-raster("model_runs/Shadow_Class_MSAVI_No_Scale_V2/products/sage_aggregate/5m_MODEL_shadow_model_and_MSAVI_V2_TILE_SM_GypsumGapNW_0101_180611.tif" )

r2_reproj<-projectRaster(from=r2, to=t, alignOnly=T)

r3<-raster("model_runs/Shadow_Class_MSAVI_No_Scale_V2/products/sage_aggregate/5m_MODEL_shadow_model_and_MSAVI_V2_TILE_SM_NaturitaNWSW_0302_180611.tif")

r3_reproj<-projectRaster(from=r3, to=t)

t2<-merge(t, r2_reproj, tolerance=0.5)

t3<-merge(t2, r3_reproj, tolerance=0.5)
```


## Moving Window
```{r}
moving_window_45m<-focal(t3, w=matrix(1/(2500*(9^2)), nrow=9, ncol=9))

writeRaster(moving_window_45m,"model_runs/Shadow_Class_MSAVI_No_Scale_V2/products/sage_aggregate/merged/moving_window_45m_window.tif")

moving_window_95m<-focal(t3, w=matrix(1/(2500*(19^2)), nrow=19, ncol=19))

writeRaster(moving_window_95m,"model_runs/Shadow_Class_MSAVI_No_Scale_V2/products/sage_aggregate/merged/moving_window_95m_window.tif")

moving_window_195m<-focal(t3, w=matrix(1/(2500*(39^2)), nrow=39, ncol=39))

writeRaster(moving_window_195m,"model_runs/Shadow_Class_MSAVI_No_Scale_V2/products/sage_aggregate/merged/moving_window_195m_window.tif")
```

